---
title: "COGS 319 Project"
output: html_notebook
---

Grid world:
A matrix of rows and colums, each square has a light value, can be occupied by an agent, and can have an obstacle
2 layers of matrix: light and obstacle + agent
light: 0 = dark, 255= max light (or 0-1)

light scaline function: 255*exp(-distance)
falls off, if we multiply d by a number less than one, drops off slower
more than 1, drops off faster

```{r}
  numRows<- 10
  numCols<- 10

makeLightGrid<- function(numRows, numCols){

  
  lightGrid<- matrix(NA, numRows, numCols)
  for(x in 1:numCols){
    for(y in 1:numRows){
      distance <- sqrt((0-y)^2+((numCols+1)/2-x)^2)
      lightGrid[y,x] <- round(255*exp(-distance*.5))
    }
  }
  return (lightGrid)
}

makeObsGrid<- function(numRows, numCols){
  obsGrid<- matrix(NA, numRows, numCols)
  obsGrid[1,]<- 1
  obsGrid[,1]<- 1
  obsGrid[numRows,]<- 1
  obsGrid[,numCols] <-1
  
    obsGrid[round(numRows/2), round(numCols/2)]<- 1
    obsGrid[round(numRows/2)+1, round(numCols/2)]<- 1
    obsGrid[round(numRows/2), round(numCols/2)+1]<- 1
    obsGrid[round(numRows/2)-1, round(numCols/2)]<- 1
    obsGrid[round(numRows/2), round(numCols/2)-1]<- 1

    obsGrid[numRows-1, round(numCols/2)]<- 0
    return(obsGrid)
    }

obsGrid<- makeObsGrid(numRows, numCols)
lightGrid<- makeLightGrid(numRows, numCols)
```


Agent traverses block matrix
blank indicates no block
1 indicates block
0 = agent is in this spot
agent cannot be on a block

Agent's variables:
cardinal: up, left, down or right pointing
light accumulator
NN
Age


functions:
(determined by NN)
turn in place to left, turn to right, step forward, step backward

move:
takes grid as argument
changes grid so agent is in new spot
return new grid

moveForward: moves in the direction bot is facing one space forward

cardinal: an ordered pair that designates next move
y, x
(rows, col)
(1,0) -> down
(0, 1) -> right
(-1,0) -> up
(0, -1) -> left
```{r}
moveForward <-function(obsGrid){
  aCur<- which(obsGrid==0, arr.ind=TRUE)
  if(is.na(obsGrid[aCur+cardinal])){
    obsGrid[aCur]<- NA
    obsGrid[aCur+cardinal]<- 0
  }
  return(obsGrid)
}
moveBackward <-function(obsGrid){
  aCur<- which(obsGrid==0, arr.ind=TRUE)
  if(is.na(obsGrid[aCur-cardinal])){
    obsGrid[aCur]<- NA
    obsGrid[aCur-cardinal]<- 0
  }
   return(obsGrid)
}

turnClock <- function(obsGrid){
  #if down, turn left
  if(all(cardinal == c(1,0))){
    cardinal <- c(0,-1)
  }
  #if left, turn up
  else if(all(cardinal == c(0,-1))){
    cardinal <- c(-1,0)
  }
  #if up, turn right
  else if(all(cardinal == c(-1,0))){
    cardinal <- c(0,1)
  }
   #if right, turn down
  else if(all(cardinal == c(0,1))){
    cardinal <- c(1,0)
  }
  return(obsGrid)
}
turnCounter <- function(obsGrid){
  #if left, turn down
  if(all(cardinal == c(0,-1))){
    cardinal <- c(1,0)
  }
  #if up turn left
  else if(all(cardinal == c(-1,0))){
    cardinal <- c(0,-1)
  }
  #if right, turn up
  else if(all(cardinal == c(0,1))){
    cardinal <- c(-1,0)
  }
   #if down, turn right
  else if(all(cardinal == c(1,0))){
    cardinal <- c(0,1)
  }
  return(obsGrid)
}

```

Test
```{r}
obsGridTest<- makeObsGrid(numRows, numCols)
lightGridTest<- makeLightGrid(numRows, numCols)
#begin facing up
cardinal <- c(-1,0)
acc<- 0
age<- 0

state <- list(
  cardinal = c(-1,0),
  obsGrid = makeObsGrid(numRows, numCols)
)

obsGridTest <- moveForward(obsGridTest)
obsGridTest <- moveForward(obsGridTest)
cardinal<- turnClock(cardinal)
obsGridTest <- moveForward(obsGridTest)

```

overall: "take turn"
loop gets current status, runs NN to see action, takes action, gets new state

(rows, col)
(1,0) -> down
(0, 1) -> right
(-1,0) -> up
(0, -1) -> left

#light function
get light takes in obsGrid, finds the position of the bot (aCur)
based on the cardinal, it takes the average of the space its in, the space in front, and the two to the right and left for the right and left sensors respectively. The sum of r and l is the number added to the accumulator
```{r}
down <- c(1,0)
right<- c(0, 1)
up<- c(-1,0)
left<- c(0, -1)

rightLight<- 0
leftLight<- 0

getlight<- function(cardinal, obsGrid, lightGrid){
  aCur<- which(obsGrid==0, arr.ind=TRUE)
  #up
  if(all(cardinal== up)){
    rightLight<- mean(lightGrid[aCur], lightGrid[aCur+cardinal], lightGrid[aCur+right], lightGrid[aCur+up+right])
      leftLight<- mean(lightGrid[aCur], lightGrid[aCur+cardinal], lightGrid[aCur+left], lightGrid[aCur+up+left])
  }
  
  #right
  else if(all(cardinal== right)){
    rightLight<- mean(lightGrid[aCur], lightGrid[aCur+cardinal], lightGrid[aCur+up], lightGrid[aCur+up+right])
      leftLight<- mean(lightGrid[aCur], lightGrid[aCur+cardinal], lightGrid[aCur+down], lightGrid[aCur+down+right])
  }
  #left
    else if(all(cardinal== left)){
    rightLight<- mean(lightGrid[aCur], lightGrid[aCur+cardinal], lightGrid[aCur+up], lightGrid[aCur+up+left])
      leftLight<- mean(lightGrid[aCur], lightGrid[aCur+cardinal], lightGrid[aCur+down], lightGrid[aCur+down+left])
    }
  #down
    else if(all(cardinal== down)){
    rightLight<- mean(lightGrid[aCur], lightGrid[aCur+cardinal], lightGrid[aCur+right], lightGrid[aCur+down+right])
      leftLight<- mean(lightGrid[aCur], lightGrid[aCur+cardinal], lightGrid[aCur+left], lightGrid[aCur+down+left])
  }
  
  acc<- acc + leftLight+rightLight
  
  return(acc)
}


```
Test for light
```{r}
obsGridTest<- makeObsGrid(numRows, numCols)
lightGridTest<- makeLightGrid(numRows, numCols)
#begin facing up
cardinal <- c(-1,0)
acc<- 0
age<- 0

obsGridTest <- moveForward(obsGridTest)
light<- getlight(obsGridTest, lightGridTest)
obsGridTest <- moveForward(obsGridTest)
light<- getlight(obsGridTest, lightGridTest)
cardinal<- turnClock(cardinal)
light<- getlight(obsGridTest, lightGridTest)
obsGridTest <- moveForward(obsGridTest)
light<- getlight(obsGridTest, lightGridTest)
accTest<- light

```


Get ir
returns the distance of the closest block


down <- c(1,0)
right<- c(0, 1)
up<- c(-1,0)
left<- c(0, -1)



```{r}

leftIR<- 0
rightIR<-0

getRightIR<- function(obsGrid){
  aCur<- which(obsGrid==0, arr.ind=TRUE)
  #up
  #checks for both Right  IR, they include the two spaces in front, sets to distance between,
  # first checks before, then to right, then up up, then up up right
  #else if because we check in order
  
  
  #up direction
  if(all(cardinal== up)){ 
    #if item in spot 1 (up), IR = distance
    if(obsGrid[aCur+up]==1){
      rightIR <- aCur[1]-aCur+up[1]/aCur[2]-aCur+up[2]
      #if item in spot 2 (right), IR = distance
    } else if (obsGrid[aCur+up+right]){
      rightIR <- aCur[1]-aCur+up+right[1]/aCur[2]-aCur+up+right[2]
      #if item in spot 3, IR= distance
    } else if (obsGrid[aCur+up+up]==1){
      rightIR <- aCur[1]-aCur+up+up[1]/aCur[2]-aCur+up+up[2]
      #if item in spot 4, IR = distance
    } else if (obsGrid[aCur+up+up+right]==1){
      rightIR <- aCur[1]-aCur+up+up+right[1]/aCur[2]-aCur+up+up+right[2]
    }}
  #down direction
  if (all(cardinal == down)){
    #spot 1
    if(obsGrid[aCur+down]==1){
         rightIR <- aCur[1]-aCur+down[1]/aCur[2]-aCur+down[2]
         #spot 2
    } else if (obsGrid[aCur+down+right]==1){
       rightIR <- aCur[1]-aCur+down+right[1]/aCur[2]-aCur+down+right[2]
       #spot 3
    } else if (obsGrid[aCur+down+down]==1){
    rightIR <- aCur[1]-aCur+down+down[1]/aCur[2]-aCur+down+down[2]
    #spot 4
    } else if (obsGrid[aCur+down+down+right]){
      rightIR <- aCur[1]-aCur+down+down+right[1]/aCur[2]-aCur+down+down+right[2]
    }
  }
  
  #right direction
  if (all(cardinal == right)){
    #spot 1
    if(obsGrid[aCur+right]==1){
         rightIR <- aCur[1]-aCur+right[1]/aCur[2]-aCur+right[2]
         #spot 2
    } else if (obsGrid[aCur+down+right]==1){
       rightIR <- aCur[1]-aCur+down+right[1]/aCur[2]-aCur+down+right[2]
       #spot 3
    } else if (obsGrid[aCur+right+right]==1){
    rightIR <- aCur[1]-aCur+right+right[1]/aCur[2]-aCur+right+right[2]
    #spot 4
    } else if (obsGrid[aCur+down+right+right]){
      rightIR <- aCur[1]-aCur+down+right+right[1]/aCur[2]-aCur+down+right+right[2]
    }
  }
  
  if (all(cardinal == left)){
    #spot 1
    if(obsGrid[aCur+left]==1){
         rightIR <- aCur[1]-aCur+left[1]/aCur[2]-aCur+left[2]
         #spot 2
    } else if (obsGrid[aCur+up+left]==1){
       rightIR <- aCur[1]-aCur+up+left[1]/aCur[2]-aCur+up+left[2]
       #spot 3
    } else if (obsGrid[aCur+left+left]==1){
    rightIR <- aCur[1]-aCur+left+left[1]/aCur[2]-aCur+left+left[2]
    #spot 4
    } else if (obsGrid[aCur+up+left+left]){
      rightIR <- aCur[1]-aCur+up+left+left[1]/aCur[2]-aCur+up+left+left[2]
    }
  }
  
  return (abs(rightIR))
}
    
    
getLeftIR<- function(obsGrid){
  aCur<- which(obsGrid==0, arr.ind=TRUE)
  if(all(cardinal== up)){ 
    #if item in spot 1 (up), IR = distance
    if(obsGrid[aCur+up]==1){
      rightIR <- aCur[1]-aCur+up[1]/aCur[2]-aCur+up[2]
      #if item in spot 2 (left), IR = distance
    } else if (obsGrid[aCur+up+leftIR]){
      rightIR <- aCur[1]-aCur+up+left[1]/aCur[2]-aCur+up+left[2]
      #if item in spot 3, IR= distance
    } else if (obsGrid[aCur+up+up]==1){
      rightIR <- aCur[1]-aCur+up+up[1]/aCur[2]-aCur+up+up[2]
      #if item in spot 4, IR = distance
    } else if (obsGrid[aCur+up+up+left]==1){
      rightIR <- aCur[1]-aCur+up+up+left[1]/aCur[2]-aCur+up+up+leftIR[2]
    }}
   if (all(cardinal == down)){
    #spot 1
    if(obsGrid[aCur+down]==1){
         rightIR <- aCur[1]-aCur+down[1]/aCur[2]-aCur+down[2]
         #spot 2
    } else if (obsGrid[aCur+down+left]==1){
       rightIR <- aCur[1]-aCur+down+left[1]/aCur[2]-aCur+down+left[2]
       #spot 3
    } else if (obsGrid[aCur+down+down]==1){
    rightIR <- aCur[1]-aCur+down+down[1]/aCur[2]-aCur+down+down[2]
    #spot 4
    } else if (obsGrid[aCur+down+down+left]){
      rightIR <- aCur[1]-aCur+down+down+left[1]/aCur[2]-aCur+down+down+left[2]
    }
   }
  #right
  if (all(cardinal == right)){
    #spot 1
    if(obsGrid[aCur+right]==1){
         rightIR <- aCur[1]-aCur+right[1]/aCur[2]-aCur+right[2]
         #spot 2
    } else if (obsGrid[aCur+up+right]==1){
       rightIR <- aCur[1]-aCur+up+right[1]/aCur[2]-aCur+up+right[2]
       #spot 3
    } else if (obsGrid[aCur+right+right]==1){
    rightIR <- aCur[1]-aCur+right+right[1]/aCur[2]-aCur+right+right[2]
    #spot 4
    } else if (obsGrid[aCur+up+right+right]){
      rightIR <- aCur[1]-aCur+up+right+right[1]/aCur[2]-aCur+up+right+right[2]
    }
  }

     if (all(cardinal == left)){
    #spot 1
    if(obsGrid[aCur+left]==1){
         rightIR <- aCur[1]-aCur+left[1]/aCur[2]-aCur+left[2]
         #spot 2
    } else if (obsGrid[aCur+down+left]==1){
       rightIR <- aCur[1]-aCur+down+left[1]/aCur[2]-aCur+down+left[2]
       #spot 3
    } else if (obsGrid[aCur+left+left]==1){
    rightIR <- aCur[1]-aCur+left+left[1]/aCur[2]-aCur+left+left[2]
    #spot 4
    } else if (obsGrid[aCur+down+left+left]){
      rightIR <- aCur[1]-aCur+down+left+left[1]/aCur[2]-aCur+down+left+left[2]
    }
  }   
 
  return(abs(leftIR))
  }
  
  
```
getBump
```{r}
bump <- 0
getBump<- function(obsGrid){
  aCur<- which(obsGrid==0, arr.ind=TRUE)
  if(is.na(obsGrid[aCur+cardinal])){
    bump<-0
  }
  else if(!is.na(obsGrid[aCur+cardinal])){
    bump<-1
  }
  

  return(bump)
}
```


NN
```{r}
clock <- 0
counter <- 0
fwd <- 0
back <- 0

input<- c(leftLight,leftIR, bump, rightIR, rightLight) 
output <- c(clock, counter, fwd, back)
weights<- matrix(NA,length(input), length(output))



```

Trial
```{r}
makeMove<- function(obsGrid){
  options<- c(moveForward, moveBackward, turnClock, turnCounter)
  action <- sample(options, 1)
  obsGrid <- action[[1]](obsGrid)
  return(obsGrid)
}



```


```{r}

runBot <- function(weights){

  # make a starting environment / state
  
  # for loop 1:n.moves
  #    accumulate light
  #    makeMove <- pass in weights, state
  #       makeMove gets sensor values
  #       makeMove calls neural net
  #       makeMove samples movement
  
  # return overall fitness
}

runGeneration <- function(genomes) {
  fitness.vals <- data.frame(bot.number=1:40, fitness=rep(0,40))
  for(i in 1:40){
    fitness <- runBot(genomes[[i]])
    fitness.vals[i, 'fitness'] <- fitness
  }
  return(fitness.vals)
  # runBot for each genome in the data frame
}

createGeneration <- function(n){
  genomes <- list()
  for(i in 1:n){
    genomes[[i]] <- matrix(runif(20, min=-1, max=1), nrow=5)    
  }
  return(genomes)
}

bots <- createGeneration(40)
bots[[1]]
```
Next steps:
  run agent through grid world
  agent that makes random decisions
  assign score to them about how well they do
  fitness component
  genetic alogrythm to evolve control structures
  
  
  Four actions: turn turn up back
  make next decision: randomly chooses one of the 4
  sample with replacement
  call function for action picked using if statement
  put in a loop, do x amount of times
  all bots in dataframe
  
  fitness score for each in dataframe
  
  
  
  
  proof of concept of evolution working: demo of modeling approach
  paper can imagine what you can learn from model hypothetically 
  
  